Estimation of Stochastic Wilson and Cowan Neural Field Model using Expectation Maximisation Algorithm

### 2011-08-17 ltx/data_paper/DataPaper.tex
edits to results section correltation analysis

where over-bar denotes complex conjugate operator. Finally, an expression for the kernel is obtained by taking the inverse Fourier tranform 

@@

By using similar arguments that were used in the derivation for the connectivity kernel support, the autocorrelation can be simplified by recognizing that

@@

In this section, a demonstration of the estimation framework for the proposed IDE neural field model is provided. To demonstrate the estimation procedure, simulated data is used to establish a ground truth for a comparison. The parameters of the observation noise and the field disturbance were assumed to be unknown, and therefore, were inferred from data. to the estimation algorithm and no assumption were made on the shape of the connectivity kernel. Data was generated employing state-space equation (\ref{eq:LS-finalformstatespacemodel}-\ref{eq:LS-QmatrixForSigmapoints}) using a square grid of $n_x= 9 \times 9$ equally spaced ($2.5$~mm) basis functions with $\sigma_{\phi}^2=1.58$~mm$^2$, assuming the observed field can be fully described by the decomposed IDE model. 100 realizations of 500 ms of data were generated and the first 100 ms was discarded allowing the model's dynamic to reach the steady state. The observation noise variance was set to 0.2. All the other parameters for the simulation were chosen as that of Chapter~\ref{LSChapter} (see Table~\ref{tab:LS-Parameters}).

\subsection{Correlation Analysis}
The correlation analysis described in \S\ref{subsec:CorrAnal} were applied to the generated data in order to approximate the unknown supports of both the connectivity kernel and the disturbance covariance function. In doing so, the unknown parameters $\sigma_{\epsilon}^2$ and $\xi$ are required and should be guessed beforehand.  To form a more accurate initial guess of the unknown parameters and explore the effect of the chosen initial guess on the estimated supports several tests were performed. First consider  \eqref{eq:EM-MMGFourier2} for estimating the disturbance covariance function. As mentioned before the power spectral densities are non-negative, therefore the minimum value of $\mathcal F \{\mathbf E\left[ R_{y_ty_t}\right] \}$ and $\mathcal F \{\mathbf E\left[ R_{y_{t+1}y_{t+1}}\right]\}$ should be chosen as the higher limit for the initial guess of $\sigma^2_{\varepsilon}$. The minimum value of the power spectral densities was found to be 0.23, Figure~\ref{fig:DisturbanceSupportExperiment} shows the support estimates for varying  $\sigma^2_{\varepsilon}$ around the calculated limit. This figure shows that for the values less than 0.23 the estimated support is close to the true one (see Figure~\ref{fig:BinaryDisturbanceWidthEstimation}(b)) while for the values above 0.23 the shape of the disturbance covariance function is distorted. The estimated support of the covariance function convolved twice with the sensor kernel for  $\sigma^2_{\varepsilon}<$0.23 is shown in Figure~\ref{fig:BinaryDisturbanceWidthEstimation} (a), where the results were normalised and thresholded for values less than 0.01. By fitting a Gaussian basis function to the result, and under the assumption that the width of the sensor, $\sigma_m^2$, was known the width of the disturbance covariance function, $\sigma_{\gamma}^2$, was analytically calculated  using \eqref{eq:AnalyticSupportCalculation}. Alternatively deconvolution can be performed numerically to directly find the support as shown in Figure~\ref{fig:DisturbanceWidthEstimationCrossSection}. The effect of the observation noise variance on the connectivity kernel support is very much like that of the disturbance covariance function, where an initial guess can be made in a similar manner. To study the influence of the synaptic parameter, $\xi=1-\frac{ T_s}{\tau}$, on the resultant support of the connectivity kernel different values of the synaptic time constant, $\tau$, in a physiologically plausible range were considered, while noting that the chosen values should be higher than the sampling time, $T_s$. In fact synaptic parameter in \eqref{eq:EM-Fourier_TF_of_Kernel}  acts as a scaling factor --- dilating or shrinking the estimated support --- since subtracting a constant in the frequency domain only affects the value at the origin in the spatial domain. The results of the correlation analysis for the connectivity kernel are shown in Figure~\ref{fig:KernelWidthEstimationExperiment} where each subplot is the normalised kernel estimate for varying synaptic time constants while holding the observation noise fixed. Note for clarification the amplitude at the origin was negated when the chosen values of $\tau$ and $\sigma^2_{\varepsilon}$ resulted in a negative value at the origin.  The result suggested that a support of approximately $\left[-5,5 \right] $ is a reasonable choice for the spatial extent of the connectivity kernel. 

>>>>

where over-bar denotes complex conjugate operator. Finally, an expression for the kernel is obtained by taking the inverse Fourier transform 

@@

By using similar arguments that were used in the derivation for the connectivity kernel support, the autocorrelation can be simplified by recognising that

@@

In this section, a demonstration of the estimation framework for the proposed IDE neural field model is provided. To demonstrate the estimation procedure, simulated data is used to establish a ground truth for a comparison. The parameters of the model that are estimated include the synaptic time constant, the shape of the connectivity kernel, the field disturbance properties and the observation noise properties. All the parameters for the simulated data given in Table~\ref{tab:LS-Parameters}. To demonstrate the estimation procedure, a Monte Carlo simulation was employed where 100 realisations of 500 ms of data were generated. The first 100 ms of each realisation was discarded allowing the model's dynamics to stabilise from the initial conditions.

\subsection{Estimation of Connectivity Kernel and Field Disturbance Support }
The correlation analysis, described in \S\ref{subsec:CorrAnal}, was applied to the simulated data to estimate the unknown support of both the connectivity kernel and the disturbance covariance function. The goal of the estimation procedure for the kernel support is to constrain to positioning and number of kernel basis functions.

From equation~\ref{eq:KernelSolution} it is clear that the an initial estimate of the synaptic time constant is required to estimate the shape of the connectivity kernel from data. \dean{Maybe we can say something like:} From equation~\ref{eq:EM-Fourier_TF_of_Kernel}, it can be seen the an error in the estimate of the value of the synaptic time constant will yield either shift up or down in the frequency domain. Since the goal of the procedure is to constrain the positioning of connectivity kernel basis functions, the initial estimate of the time constant should be conservatively \dean{higher (or lower?)}, which will result in a dilation of shape of the kernel (shift lower in the frequency domain). If the shape of the kernel estimate is wider than the actual kernel, then we can be sure that the positioning of the basis functions will capture the connectivity structure once the basis function coefficients have been estimated. If the kernel shape is too narrow, then the parameter values associated with the basis function can not compensate.

In addition to the synaptic time constant, an initial guess for the observation noise variance is also required.

In doing so, the unknown parameters $\sigma_{\epsilon}^2$ and $\xi$ are required and should be guessed beforehand. To form a more accurate initial guess of the unknown parameters and explore the effect of the chosen initial guess on the estimated supports several tests were performed. First consider \eqref{eq:EM-MMGFourier2} for estimating the disturbance covariance function. As mentioned before, the power spectral densities are non-negative. Therefore, the minimum values of $\mathcal F \{\mathbf E\left[ R_{y_ty_t}\right] \}$ and $\mathcal F \{\mathbf E\left[ R_{y_{t+1}y_{t+1}}\right]\}$ should be chosen (\dean{these aren't chosen, they are calculated from data, right? Not sure what you mean. Do mean the initial guess of `$\sigma^2_{\varepsilon}$' should be chosen such that ...? or $\sigma^2_{\varepsilon}$ is bounded by...}) as the higher limit for the initial guess of $\sigma^2_{\varepsilon}$. The minimum value of the power spectral densities was found to be 0.23.

Figure~\ref{fig:DisturbanceSupportExperiment} shows the support estimates for varying  $\sigma^2_{\varepsilon}$ around the calculated limit. This figure shows that for the values less than 0.23 the estimated support is close to the true one (see Figure~\ref{fig:BinaryDisturbanceWidthEstimation}(b)) while for the values above 0.23 the shape of the disturbance covariance function is distorted. The estimated support of the covariance function convolved twice with the sensor kernel for  $\sigma^2_{\varepsilon}<$0.23 is shown in Figure~\ref{fig:BinaryDisturbanceWidthEstimation} (a), where the results were normalised and thresholded for values less than 0.01. By fitting a Gaussian basis function to the result, and under the assumption that the width of the sensor, $\sigma_m^2$, was known the width of the disturbance covariance function, $\sigma_{\gamma}^2$, was analytically calculated  using \eqref{eq:AnalyticSupportCalculation}. Alternatively deconvolution can be performed numerically to directly find the support as shown in Figure~\ref{fig:DisturbanceWidthEstimationCrossSection}. The effect of the observation noise variance on the connectivity kernel support is very much like that of the disturbance covariance function, where an initial guess can be made in a similar manner. To study the influence of the synaptic parameter, $\xi=1-\frac{ T_s}{\tau}$, on the resultant support of the connectivity kernel different values of the synaptic time constant, $\tau$, in a physiologically plausible range were considered, while noting that the chosen values should be higher than the sampling time, $T_s$. In fact synaptic parameter in \eqref{eq:EM-Fourier_TF_of_Kernel}  acts as a scaling factor --- dilating or shrinking the estimated support --- since subtracting a constant in the frequency domain only affects the value at the origin in the spatial domain. The results of the correlation analysis for the connectivity kernel are shown in Figure~\ref{fig:KernelWidthEstimationExperiment} where each subplot is the normalised kernel estimate for varying synaptic time constants while holding the observation noise fixed. Note for clarification the amplitude at the origin was negated when the chosen values of $\tau$ and $\sigma^2_{\varepsilon}$ resulted in a negative value at the origin.  The result suggested that a support of approximately $\left[-5,5 \right] $ is a reasonable choice for the spatial extent of the connectivity kernel. 
